---
title: "Data Simulation"
---

Last step of the initialization is to create our databases inside our two slaves.

We are using 2 test dataset for testing purposes: \newline \indent
a. *winequality* for most of the prediction model;\newline \indent
b. *bladder* for our risk analysis

Remember to always run this first in order to produce the .ttl files before your SPARQL environment creation

```{r data1, eval=FALSE}
# =============================================================
# DATA PREPARATION - Prediction model
# =============================================================
setwd("~/Desktop/Federated_Learning/")
winequality <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", sep = ";")
winequality$binary_quality <- ifelse(winequality$quality >= 6, 1, 0)
colnames(winequality) <- gsub("\\.", "_", colnames(winequality))
colnames(winequality)<- tolower(colnames(winequality))

set.seed(123)
slave1_data <- winequality[sample.int(nrow(winequality), 123), ]
set.seed(321)
slave2_data <- winequality[sample.int(nrow(winequality), 321), ]
all_data <- rbind(slave1_data, slave2_data)

# convert into .ttl
library(readr)
csv_to_turtle <- function(table, output_path, slave_id) {
  df <- table
  feature_cols <- names(df)[1:11]
  y_col        <- "quality"
  
  
  lines <- c('@prefix fv: <http://example.org/fairvasc#> .\n')
  
  for (i in seq_len(nrow(df))) {
    
    uri      <- sprintf("<http://example.org/patient/%s/%04d>", slave_id, i)
    features <- paste(as.numeric(df[i, feature_cols]), collapse = " ")
    y_val    <- as.numeric(df[i, y_col])
    
    lines <- c(lines,
               sprintf('%s fv:features "%s" ;', uri, features),
               sprintf('   fv:y        "%g"  .\n', y_val)
    )
  }
  writeLines(lines, output_path)
  cat(sprintf("Written: %s  (%d rows)\n", output_path, nrow(df)))
}

csv_to_turtle(slave1_data, 'slave1.ttl', 'slave1')
csv_to_turtle(slave2_data, 'slave2.ttl', 'slave2')

# =============================================================
# DATA PREPARATION - Risk model
# =============================================================
library(survival)

bladder <- survival::bladder
bladder$treatment <- bladder$rx - 1
cox_slave1_data <- bladder[1:200, c("treatment","number","size","stop","event")]
cox_slave2_data <- bladder[201:340, c("treatment","number","size","stop","event")]
cox_all_data <- bladder

cox_to_turtle <- function(df, output_path) {
  lines <- c('@prefix : <http://example.org/> .',
             '@prefix xsd: <http://www.w3.org/2001/XMLSchema#> .')
  
  for (i in seq_len(nrow(df))) {
    lines <- c(lines,
               sprintf(':p%d :treatment %d ; :number %d ; :size %d ; :stop %d ; :event %d .',
                       i, df$treatment[i], df$number[i], df$size[i], df$stop[i], df$event[i]))
  }
  writeLines(lines, output_path)
  cat(sprintf("Written: %s  (%d rows)\n", output_path, nrow(df)))
}

cox_to_turtle(cox_slave1_data, 'cox_slave1.ttl')
cox_to_turtle(cox_slave2_data, 'cox_slave2.ttl')
```

Now the upload data for the SQL environment. Remember to create the network first and upload data with the following code right after.
```{r sql_up, eval=FALSE}
# =============================================================
# DATA LOADING - Push dataset on SQL slave nodes
# =============================================================
library(RPostgres)

SQL_TABLE_NAME     <- "federated_data"
COX_SQL_TABLE_NAME <- "cox_data"   

con_slave1 <- dbConnect(RPostgres::Postgres(), host = '127.0.0.1', port = 5433, user = 'postgres', password = 'slv1')
con_slave2 <- dbConnect(RPostgres::Postgres(), host = '127.0.0.1', port = 5434, user = 'postgres', password = 'slv2')

# --- Dataset "Prediction model" (wine quality) ---
cat("▸ Caricamento wine-quality su slave1...\n")
dbWriteTable(con_slave1, SQL_TABLE_NAME, slave1_data, overwrite = TRUE, row.names = FALSE)
cat("▸ Caricamento wine-quality su slave2...\n")
dbWriteTable(con_slave2, SQL_TABLE_NAME, slave2_data, overwrite = TRUE, row.names = FALSE)

# --- Dataset "Risk model" (bladder / Cox) ---
cat("▸ Caricamento dati Cox su slave1...\n")
dbWriteTable(con_slave1, COX_SQL_TABLE_NAME, cox_slave1_data, overwrite = TRUE, row.names = FALSE)
cat("▸ Caricamento dati Cox su slave2...\n")
dbWriteTable(con_slave2, COX_SQL_TABLE_NAME, cox_slave2_data, overwrite = TRUE, row.names = FALSE)

dbDisconnect(con_slave1)
dbDisconnect(con_slave2)

cat("SQL uplad completed.\n")

```
